\section{Algorithm}\label{sec:algorithm}

\paragraph{Purification.}
Given a QF\_NIA formula $\varphi$, the first step is \emph{purification}: every
nonlinear subterm $t$ (a product whose factors are not all constants, or an
integer division/modulo with a non-constant divisor) is replaced by a fresh
integer constant $\pure{t}$, called a \emph{pure}.
Equal subterms share the same pure.
The result $\hat\varphi$ is a QF\_LIA formula over the original variables and
the new pure constants.

\paragraph{Main loop.}
\Cref{alg:solve} shows the top-level procedure.
We maintain a set $\mathcal{A}$ of linear \emph{axioms} accumulated across iterations.
Each iteration calls a QF\_LIA solver on $\hat\varphi \land \bigwedge\mathcal{A}$.
If the LIA problem is unsatisfiable, so is the original QF\_NIA formula.
Otherwise we obtain a model $\mu$ and call \textsc{Check-Nia} (\Cref{alg:checknia}),
which returns a set $\Delta\mathcal{A}$ of new axioms.
If $\Delta\mathcal{A} = \emptyset$, the current model $\mu$ is NIA-valid and we return~$\mathit{sat}$.
Otherwise we expand $\mathcal{A}$ with $\Delta\mathcal{A}$ and repeat.

\begin{algorithm}[t]
	\caption{QF-NIA-SOLVE($\varphi$)}\label{alg:solve}
	\DontPrintSemicolon
	$\hat\varphi,\, P \leftarrow \textrm{PURIFY}(\varphi)$\;
	$\mathcal{A} \leftarrow \emptyset$\;
	\Repeat{}{
		$\mathit{res} \leftarrow \textrm{CHECK-LIA}(\hat\varphi \land \bigwedge\mathcal{A})$\;
		\lIf{$\mathit{res} = \mathit{unsat}$}{\Return $\mathit{unsat}$}
		$\mu \leftarrow \textrm{model}(\mathit{res})$\;
		$\Delta\mathcal{A} \leftarrow \textrm{CHECK-NIA}(P,\mu,\mathcal{A})$\;
		\lIf{$\Delta\mathcal{A} = \emptyset$}{\Return $\mathit{sat},\, \mu$}
		$\mathcal{A} \leftarrow \mathcal{A} \cup \Delta\mathcal{A}$\;
	}
\end{algorithm}

\begin{algorithm}[t]
	\caption{CHECK-NIA($P, \mu, \mathcal{A}$)}\label{alg:checknia}
	\DontPrintSemicolon
	$I \leftarrow \textrm{IMPLICANT}(\hat\varphi, \mu)$\;
	\lIf{$\mu \models_{\mathrm{NIA}} \varphi$}{\Return $\emptyset$}
	$R \leftarrow \{\,p \in P \mid p \in I,\;\text{literal of $I$ containing $p$ is NIA-false}\,\}$\;
	\lIf{\upshape MODEL-FIX$(\mu, R)$}{\Return $\emptyset$}
	$\Delta\mathcal{A} \leftarrow$ congruence axioms for $R$\;
	\ForEach{$p \in R$}{
		$\Delta\mathcal{A} \leftarrow \Delta\mathcal{A} \cup \textrm{AXIOMS}(p, \mu)$\;
	}
	\Return $\Delta\mathcal{A}$\;
\end{algorithm}

\paragraph{Implicant-based targeting.}
We evaluate $\hat\varphi$ under $\mu$ using \emph{three-valued NIA semantics}:
each pure $\pure{t}$ is assigned value $\mu(t)$ if its LIA model value
$\mu(\pure{t})$ agrees with its NIA value, and $\mathit{unknown}$ otherwise.
The result propagates by standard three-valued logic.
Because the LIA model already satisfies $\hat\varphi$, every literal needed
for satisfaction is either pure-free (same truth value under NIA) or
pure-containing (at worst $\mathit{unknown}$, never $\mathit{false}$), so the
overall result is either $\mathit{true}$ or $\mathit{unknown}$.
If it is $\mathit{true}$, the original variable assignment already satisfies
$\varphi$ and we return $\emptyset$ immediately.
Otherwise the traversal's cached per-literal truth values are used to extract
an \emph{implicant}: a subset $I$ of literals of $\hat\varphi$ sufficient
for $\mu \models \hat\varphi$.
We then collect into $R$ the pures appearing in literals of $I$ that evaluate
to false under NIA semantics.

\paragraph{Model repair.}
When $R \neq \emptyset$ we first attempt MODEL-FIX: a heuristic that pins
non-relevant variables and runs a small number of sub-iterations on a
restricted LIA problem, generating and accumulating axioms along the way.
If a NIA-valid model is found, \textsc{Check-Nia} returns $\emptyset$ immediately.
If not, the axioms generated during model repair are retained and we fall
through to the standard axiom generation step.

\paragraph{Axiom generation.}
For each failing pure $\pure{t} \in R$, we add to $\Delta\mathcal{A}$ every
axiom in our axiom set for $t$ that is violated by $\mu$.
Axiom classes are described in \Cref{sec:axioms}; they cover sign and zero
conditions, secant-based linear bounds for monomials $x^k$, bounds for mixed
products $x^k y^l$, tangent-plane lemmas, and integer division and modulo
constraints.
Lazy congruence axioms (requiring $\pure{t_1} = \pure{t_2}$ whenever
$t_1 = t_2$ holds in $\mu$ but $\mu(\pure{t_1}) \neq \mu(\pure{t_2})$) are
collected into $\Delta\mathcal{A}$ before per-pure axiom generation.

\paragraph{Comparison with~\cite{cimatti2018sat}.}
The overall CEGAR skeleton is the same.
The key differences are as follows.

\emph{Abstract domain.}
Cimatti et al.\ replace each product $x y$ with an uninterpreted
function symbol $f(x,y)$, yielding a QF\_UFLIA abstraction; congruence
($f(a,b) = f(a,b)$ for equal arguments) is then automatic.
We use plain fresh constants instead, so the abstract domain is QF\_LIA, and
congruence must be added as explicit axioms when violated.

\emph{Axiom generation.}
Cimatti et al.\ generate lemmas in three sequential rounds (basic axioms,
proportionality, tangent-plane), stopping as soon as a round produces at
least one lemma.
We add all violated axioms in a single pass, regardless of type.

\emph{Failing-term selection.}
Cimatti et al.\ scan all constraints of $\varphi$ falsified by $\mu$ and
collect every product $x y$ in those constraints.
We use an implicant of $\hat\varphi$ under $\mu$ and restrict to pures in
literals that fail under NIA semantics, which is more targeted.

\emph{Model repair.}
Cimatti et al.\ generate axioms and immediately return to the main LIA solve.
We first attempt MODEL-FIX: a heuristic sub-loop that pins irrelevant variables,
accumulates axioms over several restricted LIA calls, and tries to find a
NIA-valid model without returning to the outer loop.
When successful this saves an outer LIA call; when not, the axioms gathered
during the attempt are still retained and the outer loop continues as normal.

\emph{Axiom set.}
The axioms in~\cite{cimatti2018sat} cover sign, zero, neutrality,
proportionality, and tangent-plane conditions for binary products.
We retain sign, zero, and tangent-plane axioms and add secant-based bounds for
monomials $x^k$ and mixed products $x^k y^l$, described in \Cref{sec:axioms}.
For tangent-plane lemmas we optionally employ the \emph{frontier} strategy
of~\cite{cimatti2017tacas}: a per-pure bounding box tracks all model values
seen so far, and whenever a new point extends the box, extra tangent planes
are instantiated at the two mixed corners to ensure both upper and lower
bounds exist in every region.
