\documentclass[runningheads]{llncs}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepackage{amsmath,amsfonts,amssymb}
\numberwithin{table}{section}
\usepackage{stmaryrd}
\usepackage{import}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e} % ,noend
\DontPrintSemicolon%


\usepackage[hidelinks,
    pdftitle={Revisiting Incremental Linearization for Nonlinear Integer Arithmetic},
    pdfauthor={Marek Danco, Karel Chvalovsky, Mikolas Janota}]{hyperref}
\usepackage[nameinlink,capitalize,noabbrev]{cleveref} % must come after hyperref
\crefname{algocf}{Algorithm}{Algorithms}
\Crefname{algocf}{Algorithm}{Algorithms}
\renewcommand\UrlFont{\rmfamily}
\urlstyle{rm}

\newcommand{\keep}[2]{\textsf{k}(#1,#2)}
\newcommand{\sign}{\textsf{sign}}
\newcommand{\pos}{\textsf{pos}}
\newcommand{\vars}{\textsf{vars}}
\newcommand{\terms}{\textsf{terms}}
\newcommand{\lang}{\textsf{L}}
\newcommand{\dom}{\textsf{dom}}
\newcommand{\wit}{\textsf{wit}}
\newcommand{\TRUE}{\textsf{T}}
\newcommand{\FALSE}{\textsf{F}}
\newcommand{\xs}{\bar{x}}
\newcommand\napprox{\not\approx}
\newcommand{\tval}[1]{\langle{#1}\rangle}
\newcommand{\divc}[2]{#1\,\mathit{div}\,#2}
\newcommand{\modc}[2]{#1\,\mathit{mod}\,#2}
\newcommand{\solver}{\textsf{qfn2l}\xspace}

\newcommand{\model}[1]{\mu(#1)}
\newcommand{\pure}[1]{\llbracket #1 \rrbracket}

%
\title{Revisiting Incremental Linearization for Nonlinear Integer Arithmetic}
%
\author{Marek Dan\v{c}o\orcidID{0009-0008-3031-113X} \and
Karel Chvalovsk\'{y}\orcidID{0000-0002-0541-3889} \and
Mikol\'{a}\v{s} Janota\orcidID{0000-0003-3487-784X}}
\authorrunning{M. Dan\v{c}o et al.}
\institute{Czech Technical University in Prague}
%
\begin{document}
\maketitle

\begin{abstract}
	Incremental Linearization has previously been proposed for solving SMT problems over quantifier-free
	nonlinear integer arithmetic and has proven effective despite its conceptual simplicity.
	In this paper, we introduce a revised axiom set that improves convergence on polynomial constraints
	built from higher-degree monomials, such as powers and mixed products,
	a class of problems on which prior axiomatizations struggled. We present a standalone implementation
	built on top of Z3 for linear integer arithmetic and evaluate it on the NIA benchmark set from SMT-LIB.
	Our results show that the approach is competitive with state-of-the-art solvers overall and
	substantially outperforms them on benchmarks dominated by such polynomial constraints.
	\keywords{SMT \and Nonlinear Arithmetic \and NIA \and Linearization.}
\end{abstract}


\input{intro}
\input{background}
\input{algorithm}
\input{axioms}
\input{experiments}
\input{comparison}
\input{conclusion}

\begin{credits}
\subsubsection{\ackname}
The research was supported by the European Union under the project
ROBOPROX (reg.\ no.\ CZ.02.01.01/00/22\_008/0004590).
This article is part of the RICAIP project that has received funding from the
European Union's Horizon 2020 research and innovation programme under grant
agreement No~857306.
\end{credits}

\bibliographystyle{splncs04}
\bibliography{refs}

\end{document}
